---
title: "Homework 3 for Data Science, P8105 Kiana Chan"
output: github_document
---

```{r setup}
library(tidyverse)
library(dplyr)
library(readxl)
knitr::opts_chunk$set(
    echo = TRUE, 
    warning = FALSE, 
    fig.width = 8, 
    fig.height = 6, 
    out.width = "90%"
)
```

#PROBLEM 1 
* Part 1: The goal is to do some exploration of this dataset. To that end, write a short description of the dataset, noting the size and structure of the data, describing some key variables, and giving illstrative examples of observations. Then, do or answer the following (commenting on the results of each):
*Part 2: How many aisles are there, and which aisles are the most items ordered from?
*Part 3: Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.
*Part 4: Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.
*Part 5: take a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).


Load the Instacart data:

```{r}
library(p8105.datasets)
data("instacart")
```

## Part 1: Description of "Instacart" dataset: This data contains 1384617 total observations. Some key variables include order ID, product ID, order number, and product name, describing the various instacart orders. Some examples of observations in the product_name category include Bulgarian Yogurt, Spring Water, Organic Half & Half, Asparagus, etc. 


## Part 2: How many aisles are there, and which aisles are the most items ordered from?

There are 134 asiles. The aisles that most items are ordered from include: fresh vegetables, other creams cheeses, fresh fruits, packaged cheese, and canned meat seafood. 

```{r}
instacart %>% 
    group_by(aisle) %>% 
    arrange(instacart, desc(aisle))
```

count month / name observations 

```{r}
instacart %>% 
  group_by(aisle) %>%
  summarize(n_obs = n())
```



```{r}
instacart %>% 
  count(aisle, name = "n_obs")
  arrange(instacart, desc(aisle))
```
## Part 3: Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.

```{r}
instacart %>% 
  count(aisle) %>% 
  filter(n > 10000) %>% 
  mutate(aisle = fct_reorder(aisle, n)) %>% 
   ggplot(aes(x = aisle, y = n, color = aisle)) + 
    geom_point(alpha = .5) + 
    labs(
        title = "Number of Items Ordered in Each Aisle", 
        x = "Number of Items Ordered", 
        y = "Axis") + 
    
    theme(
      axis.text.x = element_text(angle = 90, hjust = 0, vjust = .35), 
      legend.position = "none" 
    )
``` 



##Part 4: Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.


```{r}
popular_items = 
instacart %>%     
    filter(aisle == c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>% 
    group_by(aisle) %>% 
    count(product_name) %>% 
    mutate(rank = rank(desc(n))) %>% 
    filter(rank < 4) %>% 
    select(-product_name) %>% 
  arrange(rank)
      knitr::kable(popular_items, 
                   caption = "Top 3 Items",
                   format = "simple", 
                   col.names = str_to_title(names(popular_items)), 
                   align = "lcc")
```

##Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).

```{r}
instacart %>%
    group_by(product_name, order_dow) %>% 
    filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>% 
    summarize(mean_hour = mean(order_hour_of_day)) %>% 
pivot_wider(
    names_from = "order_dow", 
    values_from = "mean_hour") %>% 
  
knitr::kable(
    digits = 2,
    caption = "Figure: Mean Hour of Day at Which Pink Lady Apples and Coffee Ice Cream are Ordered Each Day of Week"
)
```




#Problem 2
First, do some data cleaning:
* format the data to use appropriate variable names;
* focus on the “Overall Health” topic
* include only responses from “Excellent” to “Poor”
* organize responses as a factor taking levels ordered from “Poor” to “Excellent”

# install.packages("devtools")
```{r}
library(p8105.datasets)
data("brfss_smart2010")
```

```{r}
library(tidyverse)
library(httr)
library(jsonlite)
library(dplyr)

var_names = 
  GET("https://chronicdata.cdc.gov/views/acme-vg9e.json") %>%
  content("text") %>%
  fromJSON() %>% 
  .[["columns"]] %>% 
  .[["name"]] %>% 
  .[1:22]

brfss_smart2010 = 
  GET("https://chronicdata.cdc.gov/views/acme-vg9e/rows.json") %>% 
  content("text") %>%
  fromJSON() %>% 
  .[["data"]]

row_as_tibble = function(row_as_list, var_names) {
  var_list = row_as_list[9:30]
  names(var_list) = var_names 
  var_list[sapply(var_list, is.null)] <- NULL
  as_tibble(var_list, validate = FALSE)
}

brfss_smart2010 = 
  brfss_smart2010 %>% 
  map(.x = ., ~row_as_tibble(.x, var_names)) %>% 
  bind_rows
```
First, do some data cleaning:
* format the data to use appropriate variable names;
* focus on the “Overall Health” topic
* include only responses from “Excellent” to “Poor”
* organize responses as a factor taking levels ordered from “Poor” to “Excellent”

## Data Cleaning 
```{r}
cleaned_brfss = brfss_smart2010 %>% 
  janitor::clean_names() %>% 
  filter(topic == "Overall Health", 
         response %in% c("Excellent", "Very Good", "Good", "Poor")) %>% 
  mutate(response = as_factor(response)) %>% 
  arrange(desc(response))
```


*In 2002, which states were observed at 7 or more locations? What about in 2010?

## In 2002, states that were observed at 7 or more locations include AK, AL, CA, AZ, GA, etc. There were 49 states in total that were observed at 7 or more locations in 2002. 
```{r}
brfss_in2002 = cleaned_brfss %>%  
  rename(state = locationabbr) %>% 
  filter(year == "2002") %>% 
  group_by(state) %>% 
  count(state) %>% 
  filter(state >= 7)
```
## In 2002, states that were observed at 10 or more locations AR, AZ, AL, CA, CT, DC, etc. There were 49 states in total that were observed at 10 or more locations. 
```{r}
brfss_in2010 = cleaned_brfss %>%
  rename(state = locationabbr) %>% 
  filter(year == "2010") %>% 
    group_by(state) %>% 
    count(state) %>% 
  filter(state >= 10)
```



*Construct a dataset that is limited to Excellent responses, and contains, year, state, and a variable that averages the data_value across locations within a state. Make a “spaghetti” plot of this average value over time within a state (that is, make a plot showing a line for each state across years – the geom_line geometry and group aesthetic will help).

```{r}
brfss_new = cleaned_brfss %>% 
  mutate(data_value = as.numeric(data_value)) %>% 
  rename(state = locationabbr) %>% 
  select(year, state, response, data_value) %>% 
  filter(response == "Excellent") %>% 
  group_by(state, year) %>% 
mutate(
    ave_data_value = mean(data_value)) %>% 
ggplot(aes(x = year, y = ave_data_value, color = state)) + 
    geom_point() + 
    geom_line() + 
    labs(
      title = "Spaghetti plot average data value from 2002 to 2010",
      x = "Year",
      y = "Average Data Value"
    )
brfss_new
```


*Make a two-panel plot showing, for the years 2006, and 2010, distribution of data_value for responses (“Poor” to “Excellent”) among locations in NY State.

Note: This plot appears fine in the console section but when I knit the document, it doesn't appear the same. 

```{r}
brfss_plot = cleaned_brfss %>% 
  filter(year == "2006" | year == "2010") %>% 
  group_by(year, response)
```

```{r}
ggplot(brfss_plot, aes(x = response, y = data_value, fill = response)) +
  geom_col(position = "dodge") +
  labs(x = "Response Level", y = "Data Value", title = "Distribution of Data Value Responses Among Locations in NY State") + 
  theme_bw() + facet_grid(.~year)
  theme(
      axis.text.y = element_text(angle = 90, hjust = 0, vjust = .35), 
      legend.position = "none" 
    )

```


#Problem 3 
*Part 1: Load, tidy, and otherwise wrangle the data. Your final dataset should include all originally observed variables and values; have useful variable names; include a weekday vs weekend variable; and encode data with reasonable variable classes. Describe the resulting dataset (e.g. what variables exist, how many observations, etc).

*Part 2: Traditional analyses of accelerometer data focus on the total activity over the day. Using your tidied dataset, aggregate accross minutes to create a total activity variable for each day, and create a table showing these totals. Are any trends apparent?

*Part 3: Accelerometer data allows the inspection activity over the course of the day. Make a single-panel plot that shows the 24-hour activity time courses for each day and use color to indicate day of the week. Describe in words any patterns or conclusions you can make based on this graph.

## Question 3 Part 1

```{r}
accel_df = read_csv("./data/accel_data.csv")
cleaned_accel = accel_df %>% 
janitor::clean_names() 
```
```{r}
dataframe.test = 
cleaned_accel %>% 
pivot_longer(
  cols = activity_1:activity_1440, 
  names_to = "activity_number", 
  names_prefix = "activity_", 
  values_to = "activity_count") %>% 
  mutate(
    weekend = (day == "Saturday" | day == "Sunday"))
```


## Question 3 Part 2 
*Part 2: Traditional analyses of accelerometer data focus on the total activity over the day. Using your tidied dataset, aggregate across minutes to create a total activity variable for each day, and create a table showing these totals. Are any trends apparent?

```{r}
total_activity = 
    dataframe.test %>% 
  group_by(day_id) %>% 
  summarize(total_activities = sum(activity_count))
#knitr::kable()
total_activity %>% knitr::kable()
```

Based on the table, some trends that I can see are that the middle of the week tends to have lower levels of activity compared to the end of the week. 



## Question 3 Part 3 
*Part 3: Accelerometer data allows the inspection activity over the course of the day. Make a single-panel plot that shows the 24-hour activity time courses for each day and use color to indicate day of the week. Describe in words any patterns or conclusions you can make based on this graph.

```{r}
dataframe.test %>% 
  mutate(activity_number = as.numeric(as.numeric(activity_number))) %>% 
    group_by(day, activity_number) %>% 
    summarize(ave_value = mean(activity_count)) %>% 
    ggplot(aes(x = activity_number, y = ave_value, color = day)) + 
  geom_smooth(se = FALSE) +
    geom_line(alpha = 0.5) + 
    scale_x_discrete(limit = c(240, 480, 720, 960, 1200, 1440), 
      labels = c("4", "8", "12", "16", "20", "24")) +
    labs(
      title = "24-Hour Activity for Each Day",
      x = "Activity Number",
      y = "Activity Count"
    )
```
      
      


# Sunday around noon there is a high level of activity. In addition, you can see from the graph plot that Friday evening also has a high level ofa ctivity. This makes sense because Sunday is the weekend, and Friday evening people tend to go out. 
